Project Proposal

### 2024-11-15 CURJ_IDS702Projectproposal.qmd
Fixing research question 1 for county dataset
EDA for Q2 county data

---
title: "Project Proposal"
author: "ALL group member names here"
subtitle: "Due November 17 at 11:59pm"
format: pdf
---

@@

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
```

@@

**Research question 1:**
What is the relationship between residential segregation (continuous variable) and juvenile arrests (continuous variable)? How is this relationship affected by the county's median household income (categorical variable, where 0=below national average, 1=above national average)?

@@

```{r}
head(alabama_full)
```

@@

head(minnesota)
```

@@

```{r}
all_states_final <- all_states |>
  mutate(Region = factor(case_when(
    State == 'Texas' ~ 'West South Central', # use NA to set a value to missing
    State == 'Alabama' ~ 'East South Central',
    State == 'North Carolina' ~ 'South Atlantic',
    State == 'New York' ~ 'Middle Atlantic',
    State == 'Washington' ~ 'Pacific',
    State == 'Colorado' ~ 'Mountain',
    State == 'Minnesota' ~ 'West North Central',
    TRUE ~ State #this says to keep all other values the same
  )), School.Funding.Cat = factor(case_when(
    School.Funding.Adequacy < 0 ~ 'Inadequate', # use NA to set a value to missing
    School.Funding.Adequacy > 0  ~ 'Adequate',
  TRUE ~ NA
  )))

all_states_final[sample(nrow(all_states_final), 10), c('State', 'County', 'School.Funding.Adequacy', 'Region', 'School.Funding.Cat')]
```
```{r}

@@

```{r plots-1}

@@

@@

```{r plots-2}

@@

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?
We will meet virtually on Tuesday between 8 and 9 pm.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?
If a team member will be missing a meeting, they should provide around 10-12 hours of advance notice (the morning of).

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?
We will be using WhatsApp primarily for communication, with Slack as our secondary chat. We will send Zoom invites via email to reserve our weekly meeting time (Tuesdays 8-9 pm). Ideally, team members should respond to or acknowledge messages within 1-3 hours on weekdays and 12 hours on weekends/during the night.

>>>>

---
title: "Project Proposal"
author: "ALL group member names here"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
---

@@

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

@@

**Research question 1:** What is the relationship between residential
segregation (continuous variable) and juvenile arrests (continuous
variable)? How is this relationship affected by the region of the US
that the county / state is in (categorical variable called Region)?

@@

@@

#head(minnesota)
```

@@

```{r}
all_states_final <- all_states |>
  mutate(Region = factor(case_when(
    State == 'Texas' ~ 'West South Central', 
    State == 'Alabama' ~ 'East South Central',
    State == 'North Carolina' ~ 'South Atlantic',
    State == 'New York' ~ 'Middle Atlantic',
    State == 'Washington' ~ 'Pacific',
    State == 'Colorado' ~ 'Mountain',
    State == 'Minnesota' ~ 'West North Central',
    TRUE ~ State #this says to keep all other values the same
  )), School.Funding.Cat = factor(case_when(
    School.Funding.Adequacy < 0 ~ 'Inadequate', 
    School.Funding.Adequacy > 0  ~ 'Adequate',
  TRUE ~ NA
  ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))

all_states_final[sample(nrow(all_states_final), 10), 
c('State', 'County', 'School.Funding.Adequacy', 'Region', 
  'School.Funding.Cat')]
```

**Exploratory Plots:** 
Q1 Outcome Variable (Juvenile Arrests):
```{r plots-1}

@@

Q1 Relationship of Interest (Residential Segregation vs. Juvenile Arrests):
```{r plots-2}

@@

Q2 Outcome Variable (School Funding Adequacy):

```{r}
ggplot(all_states_final, aes(x = State, fill=School.Funding.Cat)) + 
  geom_bar() + labs(x="State", y="Count of Counties", 
  fill='School Funding Adequacy', title = "School Funding Adequacy 
  in Seven US States") + theme(axis.text.x = 
  element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_discrete(labels=c('Inadequate Funding', 
  'Adequate Funding', 'No Data Available'))
```

Q2 Relationship of Interest (School Segregation & School Funding
Adequacy):

```{r plots-4}
ggplot(all_states_final, aes(x= School.Segregation.Index, y= School.Funding.Cat))+
  geom_jitter()+ 
  labs(title = "School Segregation vs. School Funding Adequacy 
       in Counties Across 7 US States",
       x = "School Segregation Index",
       y = "School Funding Adequacy") 
```

@@

```{r plots-5}

@@

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

We will meet virtually on Tuesday between 8 and 9 pm.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

If a team member will be missing a meeting, they should provide around
10-12 hours of advance notice (the morning of).

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

We will be using WhatsApp primarily for communication, with Slack as our
secondary chat. We will send Zoom invites via email to reserve our
weekly meeting time (Tuesdays 8-9 pm). Ideally, team members should
respond to or acknowledge messages within 1-3 hours on weekdays and 12
hours on weekends/during the night.
